---
title: "GAMSTOP Licence Holders"
author: "Samuel Emerson"
date: "`r Sys.Date()`"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readxl)
library(caret)
library(RecordLinkage)
library(urltools)
library(assertive.data.uk)
library(knitr)
library(ggplot2)
library(ggpubr)

Licence_Data <- "business-licence-register.xlsx" %>% excel_sheets() %>% 
  set_names() %>% map(read_excel,path = "business-licence-register.xlsx")
```

## Problem Outline

* Since the 31st of March 2020, all online gambling Licence holders must be part of the GAMSTOP scheme. 
* Therefore, we required a comprehensive list of all Licence holders in the UK which are subject to integration with GAMSTOP.
* We assume from now on that a Licence holder can be identified by their account number.

## Licence Criteria

The criteria for integration with GAMSTOP can be defined by the following 4 key points: 
 1. A licence holder has a licence which is remote.
 2. A licence holder has a licence whose status is active.
 3. A licence holder has a licence whose activities include one or more of: Casino, Bingo, External Lottery Manager, General Betting Standard – Virtual Event, General Betting Standard – Real Event and Pool Betting.
 4. The licence holder has a UK domain.

## Licences - Remote Check

* The criteria we use for determining if a licence is remote is if the type is remote or the licence number contains the letter "R".

* If the two criteria disagree we flag this in an automated process. We can look at the criteria confusion matrix to visually check however:

```{r cars, echo = FALSE}
ConMat <- as.table(confusionMatrix(as.factor(Licence_Data$Licences$Type=="Remote"),
                as.factor(grepl("R",Licence_Data$Licences$`Licence Number`))))
rownames(ConMat) <- c("R not in No.","R in No.")
colnames(ConMat) <- c("Other Type","Remote Type")
kable(ConMat)
```

## Licences - Active Check

* The criteria we use for determining if a licence is active is if the status is active or if the licence has a start date by no end date.

* If the two criteria disagree we again flag this in an automated process. The criteria confusion matrix will show if any licenses have this behavior:

```{r pressure, echo = FALSE}
ConMat.2 <- as.table(confusionMatrix(as.factor(Licence_Data$Licences$Status=="Active"),
                as.factor(!is.na(Licence_Data$Licences$`Start Date`) & is.na(Licence_Data$Licences$`End Date`))))
rownames(ConMat.2) <- c("Other Status","Active Status")
colnames(ConMat.2) <- c("No Start or End","Start and No End")
kable(ConMat.2)
```

## Licences - Account Errors

* As well as checking the specified criteria, we must also ensure that account numbers match the first 6 digits on their corresponding licence number.

* This check will also be flagged if it fails, and as before we can visually check this through a frequency table:

```{r aa, echo = FALSE}
Mat <- table(factor(Licence_Data$Licences$`Account Number`==as.numeric(substr(Licence_Data$Licences$`Licence Number`,1,6)),levels = c(TRUE,FALSE)))
kable(Mat,col.names = c("Match","Frequency"))
```

## Potential Licences

* All information surrounding licences is contained in one data frame, and so using the checks we have defined we can filter out licences which do not meet the criteria.

```{r ab, echo = FALSE}
GAMSTOP_Data <- Licence_Data$Licences %>% 
  add_column(Flag_Remote = rep(FALSE,nrow(Licence_Data$Licences))) %>%
  add_column(Flag_Active = rep(FALSE,nrow(Licence_Data$Licences))) %>%
  add_column(Flag_Account = rep(FALSE,nrow(Licence_Data$Licences)))

remote_names <- unique(GAMSTOP_Data$Type)[which(sapply(unique(GAMSTOP_Data$Type),
                                                       FUN = function(u){levenshteinDist(u,"Remote")})<2)]
GAMSTOP_Data <- GAMSTOP_Data %>% 
  filter(Type%in%remote_names | grepl("R",`Licence Number`)) %>%
  mutate(Flag_Remote=(!Type%in%remote_names | !grepl("R",`Licence Number`)))

active_names <- unique(GAMSTOP_Data$Status)[which(sapply(unique(GAMSTOP_Data$Status),
                                                         FUN = function(u){levenshteinDist(u,"Active")})<2)]
GAMSTOP_Data <- GAMSTOP_Data %>% 
  filter(Status%in%active_names | (!is.na(`Start Date`) & is.na(`End Date`))) %>%
  mutate(Flag_Active=(!Status%in%active_names | (is.na(`Start Date`) | !is.na(`End Date`))))

activity_names <- unique(GAMSTOP_Data$Activity)[unique(which(outer(unique(GAMSTOP_Data$Activity),
                                                                   c("Casino","Bingo","External Lottery Manager",
                                                                     "General Betting Standard - Virtual Event",
                                                                     "General Betting Standard - Real Event",
                                                                     "Pool Betting"),FUN = levenshteinDist)<2,arr.ind = T)[,1])]
GAMSTOP_Data <- GAMSTOP_Data %>% filter(Activity%in%activity_names)

GAMSTOP_Data <- GAMSTOP_Data %>% 
  mutate(Flag_Account = `Account Number`!=as.numeric(substr(`Licence Number`,1,6)))

kable(head(GAMSTOP_Data))
```

## Accounts - UK Domains

* With the licences now meeting three of the four criteria we can now consider the licence holders attached to these licences.

* If the licence holder has a UK domain and an licence covered by the criteria then they should be integrated with GAMSTOP.

* Defining whether a licence holder has a UK domain is not a trivial task however...

## Accounts - Domain Checks

In order to determine if a licence holder has a UK domain, we detail 4 checks (with a licence holder only needing to pass one):

 1. The business name of the licence holder contains the acronym "UK".
 2. The trading name of the licence holder contains the acronym "UK".
 3. The suffix of the domain name of the one of the licence holders websites contains ".uk".
 4. The licence holders head office address either has a UK postcode or has a listed country which is one of "Alderney", "Great Britain", "Guernsey", "Isle of Man", "Northern Ireland" or "United Kingdom".
 
## Domain Check Breakdown

* The following charts gives a breakdown of the domain checks for the licences we previously obtained:

```{r ac, echo = FALSE}
Business_Data <- Licence_Data$Businesses %>% 
  filter(grepl("UK | uk | Uk",`Licence Account Name`))
                                                    
GAMSTOP_Data <- left_join(GAMSTOP_Data,Business_Data,by = "Account Number",
                          multiple = "all")

Trading_Data <- Licence_Data$TradingNames %>% 
  filter(Status=="Active" & grepl("UK | uk | Uk",`Trading Name`)) %>%
  select(`Account Number`,`Trading Name`)

GAMSTOP_Data <- left_join(GAMSTOP_Data,Trading_Data,by = "Account Number",
                          multiple = "all")

Domain_Data <- Licence_Data$DomainNames %>% 
  filter(Status=="Active" & grepl("uk",suffix_extract(domain(`Domain Name`))$suffix)) %>%
  select(`Account Number`,`Domain Name`)

GAMSTOP_Data <- left_join(GAMSTOP_Data,Domain_Data,by = "Account Number",
                          multiple="all")

uk_names <- unique(GAMSTOP_Data$Type)[which(sapply(unique(GAMSTOP_Data$Type),
                                                       FUN = function(u){levenshteinDist(u,"Remote")})<2)]

Address_Data <- Licence_Data$Addresses %>%
  filter(Country%in%c("Alderney","Great Britain","Guernsey","Isle of Man",
                      "Northern Ireland","United Kingdom") | is_uk_postcode(Postcode)) %>%
  select(`Account Number`,Country)

GAMSTOP_Data <- left_join(GAMSTOP_Data,Address_Data,by = "Account Number",
                          multiple="all")

GAMSTOP_unique <- GAMSTOP_Data[which(!duplicated(GAMSTOP_Data$`Account Number`)),]
GAMSTOP_breakdown <- !is.na(GAMSTOP_Data[,(ncol(GAMSTOP_unique)-3):ncol(GAMSTOP_unique)])
GAMSTOP_bar_1 <- as.data.frame(table(apply(GAMSTOP_breakdown,1,sum)))

bar_1 <- ggplot(GAMSTOP_bar_1, aes(x=Var1, y=Freq, fill=Var1)) +
  geom_bar(stat="identity", width=1, color="white") +
  theme(legend.position="none") +
  scale_fill_brewer(palette="RdYlGn") + xlab("Checks Passed") + 
  ylab("Accounts")

GAMSTOP_breakdown_2 <- GAMSTOP_breakdown
for(i in c("Business Name","Trading Name","Website Name","Head Office")){
  j <- which(c("Business Name","Trading Name","Website Name","Head Office")==i)
  GAMSTOP_breakdown_2[which(GAMSTOP_breakdown[,j]),j] <- i
}
GAMSTOP_breakdown_2 <- c(GAMSTOP_breakdown_2)
GAMSTOP_breakdown_2 <- GAMSTOP_breakdown_2[-which(GAMSTOP_breakdown_2=="FALSE")]

GAMSTOP_bar_2 <- as.data.frame(table(GAMSTOP_breakdown_2))

bar_2 <- ggplot(GAMSTOP_bar_2, aes(x=GAMSTOP_breakdown_2, y=Freq, fill=GAMSTOP_breakdown_2)) +
  geom_bar(stat="identity", width=1, color="white") +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Dark2") + xlab("Check Type") + 
  ylab("Accounts")

ggarrange(bar_1,bar_2,ncol=1,nrow=2)
```

## Summary & Next Steps

* Combining the information we have on domains with the information we have on licences then gives the desired list of operators which should be integrated with GAMSTOP.

* When checking the licence information we only encountered one flag regarding whether that licence was active, but this licence holder showed no evidence of having a UK domain.

* As further work, a rigorous definition of what constitutes a UK domain would be beneficial for a more accurate assessment of which operators need integrating. This could be achieved through liaison with the operators as well as the gambling commission.
